---
title: "Session 3 : Topic model"
author: "cb,jcrm,bc,oc"
date: "`r Sys.Date()`"
output: 
  html_document :
    toc: true
    toc_float: true
    toc_depth: 3
---

# Tools

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=TRUE, message=FALSE, warning=FALSE)
#from session 1 & 2
library(tidyverse)
library(udpipe)
library(flextable)
library(cowplot)
library(quanteda)
library(quanteda.textmodels)
library(quanteda.textstats)
library(quanteda.textplots)
library(syuzhet)  #analyse du sentimeent

# new for session 3
library(FactoMineR)
library(factoextra)
library(igraph)
library(ggwordcloud)
library(ggrepel)

library(Rtsne)
library(tidytext)

theme_set(theme_minimal())

t1=Sys.time()

```

# The basic LDA

proposed by @blei

## a few of theory

## Application

# some other models

Topics model are popular, and quicly new models were suggested. 

## STM

Introduce de control a third variable, as a regression approach model, very usefull to introduce a time parameter. But also let the topic being correlateed, which open ti hierarchical topics

## semi-supervised

We could help the solution by predefining key tokens. Seed_lda

# the optimal number of topic

No optimal analytics methods

but some systematic approach by comparing index for different solutions. 


# Notes

Beware to computing time! Best to sample for testing. (come back to the beginning)

```{r 20}
t2=Sys.time()
t<- t2-t1
print(t)
```


See you to later and [go to session 4](https://benaventc.github.io/NLP_lecture_PSLWeek/session4.html)

Some exercises before, for training :

* Explore the adjectives world
* Compare administratiion (main) in term of qualificatives.

# References

